---
title: "Using the minimaxApprox package"
author: "Avraham Adler"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Using the minimaxApprox package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [Weirstrass Approximation Theorem](https://mathworld.wolfram.com/WeierstrassApproximationTheorem.html)
lets us approximate any function by polynomials mathematically to any desired
degree of accuracy. This is important in the computer age as there are many
functions which are difficult to properly represent in code, especially given
the limitations of floating-point representation
([Goldberg 1991](https://docs.oracle.com/cd/E19957-01/806-3568/ncg_goldberg.html)).
Therefore, it is common to use polynomial approximations to such functions,
especially near poles or singularities.

There are various kinds of polynomial approximation, including Maclaurin and
Taylor series. 

 * Chebyshev

 * Minimax

 * Rational
 
 * Rational Minimax
 
## Description of Process

 * step through with some graphs
 
## Usage Examples

## Failure modes

## Future Research

## References




## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
