---
title: "Using the minimaxApprox package"
author: "Avraham Adler"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Using the minimaxApprox package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [Weirstrass Approximation Theorem](https://mathworld.wolfram.com/WeierstrassApproximationTheorem.html)
lets us approximate any function by polynomials mathematically to any desired
degree of accuracy. This is important in the computer age as there are many
functions which are difficult to properly represent in code, especially given
the limitations of floating-point representation
([Goldberg 1991](https://docs.oracle.com/cd/E19957-01/806-3568/ncg_goldberg.html)).
Therefore, it is common to use approximations to such functions, especially near
poles or singularities.

There are various kinds of approximations. Perhaps the simplest and best known
are the polynomial approximations. Most people are introduced to them through
Maclaurin and Taylor series. These are polynomial approximations to the function
around zero (Maclaurin) or around a point other than zero (Taylor). In some
ways, they are the *best* polynomial approximation at that point.

There are better approximations than a pure polynomial, though. Henri Padé
showed that the ratio of two polynomials---a rational function---can provide
better estimates to the function than a polynomial can.



 * Chebyshev

 * Minimax

 * Rational
 
 * Rational Minimax
 
## Description of Process

 * step through with some graphs
 
## Usage Examples

## Failure modes

## Future Research

## References




## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```


## References
